Parallel global optimization algorithm with uniform convergence for solving a set of constrained global optimization problems\thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}

### 2020-06-15 2020/OPTIMA/optima_paper.tex
minor corrections

\section{Introduction}
\label{sec-intro}
Global optimization problems with non-convex constraints are considered traditionally to be one
of the most difficult optimization problems.
 The finding of the global minimum of a function of several variables appears to be more
difficult often than a local optimization in a thousand-dimensional space.
For the latter, application of the simplest gradient descent method may appear to be enough,
whereas to \textit{guarantee} the finding of the global optimum the optimization
methods have to accumulate the information on the behaviour of the objective function in the
whole search domain \cite{Jones2009,Paulavicius2011,Evtushenko2013,Strongin2000}.
The solving of a series of such problems at limited computational resources is even more
difficult problem: besides the search of the global extremum, it is necessary to distribute the
computational resources in such a way that the location of the global extremum would be
estimated with approximately equal quality in all problems being solved at once.
Usually, a series of \(q\) problems is solved either sequentially or in parallel by portions of \(p\ll
q\) problems where \(p\) is a number of parallel computational devices.
Such an approach leads to a situation when at every time moment until the stop of computations
some problems remain, in which the estimate of the global optimum is not obtained yet whereas
in the problems from the beginning of the list the optimum may be estimated even with excess
precision.

@@

There are various methods solving the considered multidimensional problem directly
\cite{SergeyevKvasov2017, Jones2009} as well as efficient methods for solving the univariate problems \cite{Norkin1992, Strongin2000}.
In the present work, a one-dimensional method is considered, which is applied together with the
dimensionality reduction scheme.
The use of space-filling curve (or \textit{evolvent}) $y(x)$, where
\begin{equation}
\label{cube}
\lbrace y\in \mathbf{R}^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant
N\rbrace=\{y(x):0\leqslant x\leqslant 1\},
\end{equation}
 is a classical scheme of reduction of the
dimensionality of the initial problem for the global optimization algorithms
\cite{Sergeyev2013}.
A mapping of the type (\ref{cube}) allows reducing a multidimensional problem to a univariate one at the expense of worsening its properties. In particular, the
one-dimensional function \(\varphi(y(x))\) is not lipschitzian but h\"{o}lderian one:
\begin{displaymath}
\label{holder}
|\varphi(y(x_1))-\varphi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}}, \; x_1,x_2\in[0,1],
\end{displaymath}
where the H\"{o}lder constant \(H\) is related to the Lipschitz one \(L\) by the relation
\begin{displaymath}
  H=2L\sqrt{N+3}.
\end{displaymath}

@@

The index algorithm of global search (IAGS) for solving the one-dimensional problems
%(\ref{eq:constrained_problem})
considered here implies the constructing of a sequence of points
\(x_k\), which the values of the minimized function or of the constraints \(z_k = g_s(x_k)\) are
computed in.
To account for the latter, the index scheme \cite{Strongin2000} was used.
Let \(Q_0=[0,1]\). The constraint having the index \(j\) is satisfied in all points of the domain
\begin{displaymath}
  Q_j=\left\{x\in [0,1]:g_j(x)\leq 0\right\},
\end{displaymath}
which is called feasible for this constraint.
At that, the feasible domain \(D\) of the initial problem is defined by the equality
 \(D=\cap _{j=0}^{m}Q_{j}\).
A trial in the point \(x\in [0,1]\) consists in a sequence computing of the values
\(g_{1}(x),...,g_{\nu }(x)\)
where the values of the indices \(\nu\) are defined by the conditions:
\(x\in Q_{j},0\leqslant j<\nu ,x\notin Q_{\nu }\).
The finding of the first violation of any constraint terminates the trial at the point \(x\).
In the case when the point \(x\) is feasible, i.e. \(x\in D\), the trial includes the computing of all
functions of the problem.
At that, the value of the index is accepted to be equal to the magnitude \(\nu =m+1\).
The pair \(\nu =\nu (x), z=g_{\nu }(x)\) where the index \(\nu\) falls into the borders \(1\leqslant
\nu \leqslant m+1\) is called the result of the trial at the point \(x\).

Such an approach to conducting the trials allows reducing the initial problem with functional
constraints to an unconstrained problem of minimization of a discontinuous function:

\begin{displaymath}
  \begin{array}{lr}
    \psi (x^{*})=\min_{x\in [0,1]}\psi (x), \\
    \psi (x)={\begin{cases}g_{\nu }(x)/H_{\nu },&\nu <M,\\(g_{M}(x)-g_{M}^{*})/H_{M},&\nu
=M.\end{cases}}
  \end{array}
\end{displaymath}
Here \(M=\max_{}^{}\left\{\nu (x):x\in [0,1]\right\}\) and \(g_{M}^{*}=\min
_{}^{}\left\{g_{M}(x):x\in \cap _{i=0}^{M-1}Q_{i}\right\}\).
Because of the definition of the number \(M\), the problem of finding \(g_{M}^{*}\) always
has a solution. If \(M=m+1\), then \(g_{M}^{*}=\varphi(x^{*})\).
The function \(\psi (x)\) satisfies the H\"{o}lder condition within the sets
\(\cap _{i=0}^{j}Q_{i},0\leq j\leq M-1\)
with the constant 1, and \(\psi (x)\) can have jump discontinuities at the boundaries
of these sets.
Despite the values of the H\"{o}lder constants \(H_k\) and the value \(g_{M}^{*}\) are not
known in advance, these ones can be estimated in the course of solving the problem.
A set of triples \(\{(x_k,\nu_k,z_k)\}, 1\leqslant k\leqslant n\), constitutes the search information
accumulated by the method upon executing \(n\) steps.

@@

Step 4. For all nonempty sets \(I_{\nu },\nu ={\overline {1,M}}\) compute the estimates
\begin{equation}
  \label{eq:step_4}
  z_{\nu }^{*}={\begin{cases}\min\{g_{\nu }(x_{i}):x_{i}\in I_{\nu }\},&\nu =M,\\
  0,&\nu <M,\end{cases}}.
\end{equation}

@@

Step 7. Perform the next trial in the middle of the interval \((x_{t-1},x_{t})\) if the indices of its
end points are not the same: \(x^{k+1}={\frac {1}{2}}(x_{t}+x_{t-1})\).
Otherwise, perform the trial at the point
\begin{displaymath}
  x^{k+1}={\frac {1}{2}}(x_{t}+x_{t-1})-\operatorname {sgn}(z_{t}-z_{t-1}){\frac {|z_{t}-
z_{t-1}|^{n}}{2r_{\nu }\mu _{\nu }^{n}}},\nu =\nu (x_{t})=\nu (x_{t-1}),
\end{displaymath}
and then increment \(k\) by 1.

@@

The conditions of convergence of the described in Section \ref{sec:method} method in case of \(q=1\) are given in \cite{Strongin2000}.
\begin{theorem} (Sufficient convergence conditions)
  \label{th:single_conv}
  Assume that the following conditions are true:
  \begin{enumerate}
    \item The problem (\ref{eq:constrained_problem}) has a solution.
    \item Functions \(g_j(y)\leqslant 0, 1\leqslant j\leqslant m + 1\), are Lipschitzian with
respective constants \(L_i\) over the domain \(D\) (here \(g_{m+1}(y)=\varphi(y)\)).
    \item Beginning with some step (i.e. if \(k\) from (\ref{eq:points}) is sufficiently large),
    the values \(\mu_\nu\) from (\ref{step2}) satisfy the inequalities
    \begin{equation}
      r_\nu\mu_\nu > 2^{3-1/N}L_\nu \sqrt{N+3},\: 1\leqslant \nu \leqslant m + 1.
    \end{equation}
  \end{enumerate}
  Then any limit point \(\overline{y}\) of the sequence \(\{y_k\} = \{y(x_k)\}\) generated
  by the index method while solving the problerm
  (\ref{eq:constrained_problem}) is admissible and satisfies the conditions
\begin{equation}
  \label{eq:conv_cond}
  \varphi(\overline{y})=\{ \varphi(y^k): g_i(y^k)\leqslant 0,1\leqslant j\leqslant m, k=1,2,\dots\}=\varphi(y^*),
\end{equation}
  where \(y^*\) from (\ref{eq:constrained_problem}).
\end{theorem}

@@

\begin{theorem} (On convergence conditions) Let the conditions of the Theorem \ref{th:single_conv}
  are satisfied for all of \(q\) problems from the set (\ref{eq:many_problems}) separately.
  Then solving all of the \(q\) problems by the algorithm from Section \ref{sec:method} will
  generate \(q\) infinite sequences, for the limit points of which the Theorem \ref{th:single_conv}
  will be also true.
\end{theorem}
\begin{proof}
  Let us consider two random problems from the set (\ref{eq:many_problems}).
  \begin{equation}
      \begin{array}{lr}
        \min\{\varphi(y):y\in D_1,\: g_j^\varphi(y)\leqslant 0, 1\leqslant j\leqslant m_1\} \\
        \min\{\psi(y):y\in D_2,\: g_j^\psi(y)\leqslant 0, 1\leqslant j\leqslant m_2\}
      \end{array}
  \end{equation}
  Let us denote the characteristics (\ref{step3_1}) for the first problem as \(R_\varphi(i)\)
  and for the second problem as \(R_\psi(j)\). Considering that, we have:
  \begin{equation}
      \begin{array}{lr}
        R_\varphi(t_\varphi)=\max_{1\leqslant i\leqslant k}R_\varphi(i), \\
        R_\psi(t_\psi)=\max_{1\leqslant j\leqslant s}R_\psi(j),
      \end{array}
  \end{equation}
  where \(k\) corresponds to the number of trials in the first problem and \(s\)
  corresponds to the number of trials in the second one. The sequence of trials \(\{v^k\}\)
  corresponds to the first problem and the sequence of trials \(\{u^s\}\)
  corresponds to the second problem. The values \(z^k=g^\varphi_\nu(v_k),\nu =\nu (v_{k})\) correspond to the
  trial points \(\{v^k\}\) and the values \(w^s=g^\psi_\nu(u_s),\nu =\nu(u_{s})\) correspond to the
  trial points \(\{u^s\}\).

@@

  Let us consider the first case: from (\ref{step2})
  \begin{displaymath}
    \frac{|z_i-z_{i-1}|}{\Delta_i} \leqslant \mu_{\varphi,\nu}.
  \end{displaymath}
  Considering that, we can establish an upper bound:
  \begin{displaymath}
    \frac{(z_i-z_{i-1})^2}{r_\nu^2\mu_{\varphi,\nu}^2\Delta_i}=\frac{(z_i-z_{i-1})^2\Delta_i}{(r_\nu\mu_{\varphi,\nu}\Delta_i)^2}
    \leqslant \frac{\Delta_i}{r_\nu^2}.
  \end{displaymath}
  Therefore using the first branch form rule (\ref{step3_1}) we have inequality
  \begin{equation}
    \label{eq:th1}
    R_\varphi(i)\leqslant\Delta_i(1 + \frac{1}{r_\nu^2}) - \frac{2(z_i+z_{i-1}-2z^*_\nu)}{r_\nu\mu_{\varphi,\nu}}.
  \end{equation}
  Because \(\overline{v}\) is the limit point of the sequence \(\{v^k\}\) and \(\varphi(y(\overline{v}))\leqslant z^*_{\nu}, \nu=m+1\) or
  \(z^*_\nu=0, \nu<m+1\) and \(z_{i-1},z_i\to 0\) at \(k\to\infty\):
  \begin{equation}
    \label{eq:th2}
    \Delta_i\to 0, z_i+z_{i-1} - 2 z_\nu^*\to 0.
  \end{equation}
  In the second case (when one of the other two branches of the rule (\ref{step3_1}) is applied) we have
  \begin{displaymath}
    R_\varphi(i)=2\Delta_i - 4\frac{z_i-z^*_\nu}{r_\nu\mu_{\varphi,\nu}}
  \end{displaymath}
  If \(z^*_\nu \ne 0\) then \(z_i-z^*_\nu \geqslant 0\) and
  \begin{equation}
    \label{eq:th3}
    R_\varphi(i)=2\Delta_i - 4\frac{z_i-z^*_\nu}{r_\nu\mu_{\varphi,\nu}} \leqslant 2\Delta_i.
  \end{equation}
  Otherwise because of \(\overline{v}\) is a feasible point \(z_i\to 0\) at \(k\to\infty\).

@@

  If \(\nu(w_{j-1})=\nu(w_{j})=\alpha\) then the first branch of the rule (\ref{step3_1}) is applied and
  \begin{dmath}
    R_\psi(j)=\Delta_j + \frac{(w_j-w_{j-1})^2}{r_\alpha^2\mu_{\psi,\alpha}^2\Delta_j}
      - \frac{2(w_j+w_{j-1}-2w^*_\alpha)}{r_\alpha\mu_{\psi,\alpha}} \geqslant
      \Delta_j - \frac{2(w_j+w_{j-1}-2w^*_\alpha)}{r_\alpha\mu_{\psi,\alpha}} =
      \Delta_j - \frac{2\Delta_j(w_{j-1}-w_j)}{r_\alpha\mu_{\psi,\alpha}\Delta_j} \geqslant
      \Delta_j - \frac{\Delta_j}{r_\alpha} = \Delta_j\left(1-\frac{2}{r_\alpha}\right).
    \label{eq:th4}
  \end{dmath}

>>>>

\section{Introduction}
\label{sec-intro}
Global optimization problems with non-convex constraints are considered to be one
of the most difficult optimization problems.
Finding of the global minimum of a function of several variables appears to be more
difficult often than a local optimization in a thousand-dimensional space.
For the latter, application of the simplest gradient descent method may appear to be enough,
whereas to \textit{guarantee} the finding of the global optimum the optimization
methods have to accumulate the information on the behaviour of the objective function in the
whole search domain \cite{Jones2009,Paulavicius2011,Evtushenko2013,Strongin2000}.
The solving of a series of such problems at limited computational resources is even more
difficult problem: besides the search of the global extremum, it is necessary to distribute the
computational resources in such a way that the location of the global extremum would be
estimated with approximately equal quality in all problems being solved at once.
Usually, a series of \(q\) problems is solved either sequentially or in parallel by portions of \(p\ll
q\) problems where \(p\) is a number of parallel computational devices.
Such an approach leads to a situation when at every time moment until the stop of computations
some problems remain, in which the estimate of the global optimum is not obtained yet whereas
in the problems from the beginning of the list the optimum may be estimated even with excess
precision.

@@

There are various methods solving the considered multidimensional problem directly
\cite{SergeyevKvasov2017, Jones2009} as well as efficient methods for solving the univariate problems \cite{Norkin1992, Strongin2000}.
In the present work, a one-dimensional method is considered, which is applied together with the
dimensionality reduction scheme.
The use of space-filling curve (or \textit{evolvent}) $y(x)$, where
\begin{equation}
\label{cube}
\lbrace y\in \mathbf{R}^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant
N\rbrace=\{y(x):0\leqslant x\leqslant 1\},
\end{equation}
 is the well-known scheme of dimensionality reduction
 of the initial problem for the global optimization algorithms
\cite{Sergeyev2013}.
A mapping of the type (\ref{cube}) allows reducing a multidimensional problem to a univariate one at the expense of worsening its properties. In particular, the
one-dimensional function \(\varphi(y(x))\) is not lipschitzian but h\"{o}lderian one:
\begin{displaymath}
\label{holder}
|\varphi(y(x_1))-\varphi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}}, \; x_1,x_2\in[0,1],
\end{displaymath}
where the H\"{o}lder constant \(H\) is related to the Lipschitz one \(L\) by the relation
\begin{displaymath}
  H=2L\sqrt{N+3}.
\end{displaymath}

@@

The index algorithm of global search (IAGS) for solving the one-dimensional problems
%(\ref{eq:constrained_problem})
considered here implies the constructing of a sequence of points
\(x_k\), which the values of the minimized function or of the constraints \(z_k = g_s(x_k)\) are
computed in.
To account for the latter, the index scheme \cite{Strongin2000} was used.
Let \(Q_0=[0,1]\). The constraint having the index \(j\) is satisfied in all points of the domain
\begin{displaymath}
  Q_j=\left\{x\in [0,1]:g_j(x)\leq 0\right\},
\end{displaymath}
which is called feasible for this constraint.
At that, the feasible domain \(D\) of the initial problem is defined by the equality
 \(D=\cap _{j=0}^{m}Q_{j}\).
A trial in the point \(x\in [0,1]\) consists in a sequence computing of the values
\(g_{1}(x),...,g_{\nu }(x)\)
where the values of the indices \(\nu\) are defined by the conditions:
\(x\in Q_{j},0\leqslant j<\nu ,x\notin Q_{\nu }\).
The first observed violation of any constraint terminates the trial at the point \(x\).
In the case when the point \(x\) is feasible, i.e. \(x\in D\), the trial includes the computing of all
functions of the problem.
At that, the value of the index is accepted as \(\nu =m+1\).
The pair \(\nu =\nu (x), z=g_{\nu }(x)\) where the index \(\nu\) falls into the borders \(1\leqslant
\nu \leqslant m+1\) is called the result of the trial at the point \(x\).

Such an approach to conducting the trials allows reducing the initial problem with functional
constraints to an unconstrained problem of minimization of a discontinuous function:
\begin{displaymath}
  \begin{array}{lr}
    \psi (x^{*})=\min_{x\in [0,1]}\psi (x), \\
    \psi (x)={\begin{cases}g_{\nu }(x)/H_{\nu },&\nu <M,\\(g_{M}(x)-g_{M}^{*})/H_{M},&\nu
=M.\end{cases}}
  \end{array}
\end{displaymath}
Here \(M=\max_{}^{}\left\{\nu (x):x\in [0,1]\right\}\) and \(g_{M}^{*}=\min
_{}^{}\left\{g_{M}(x):x\in \cap _{i=0}^{M-1}Q_{i}\right\}\).
Because of the definition of the number \(M\), the problem of finding \(g_{M}^{*}\) always
has a solution. If \(M=m+1\), then \(g_{M}^{*}=\varphi(x^{*})\).
The function \(\psi (x)\) satisfies the H\"{o}lder condition within the sets
\(\cap _{i=0}^{j}Q_{i},0\leq j\leq M-1\)
with the constant 1, and \(\psi (x)\) can have jump discontinuities at the boundaries
of these sets.
Despite the values of the H\"{o}lder constants \(H_k\) and the value \(g_{M}^{*}\) are not
known in advance, these ones can be estimated in the course of solving the problem.
A set of triples \(\{(x_k,\nu_k,z_k)\}, 1\leqslant k\leqslant n\), constitutes the search information
accumulated by the method upon executing \(n\) steps.

@@

Step 4. For all nonempty sets \(I_{\nu },\nu ={\overline {1,M}}\) compute the estimates
\begin{equation}
  \label{eq:step_4}
  z_{\nu }^{*}={\begin{cases}\min\{g_{\nu }(x_{i}):x_{i}\in I_{\nu }\},&\nu =M,\\
  0,&\nu <M.\end{cases}}
\end{equation}

@@

Step 7. Perform the next trial in the middle of the interval \((x_{t-1},x_{t})\) if the indices of its
end points are not the same: \(x^{k+1}={\frac {1}{2}}(x_{t}+x_{t-1})\).
Otherwise, perform the trial at the point
\begin{displaymath}
  x^{k+1}={\frac {1}{2}}(x_{t}+x_{t-1})-\operatorname {sgn}(z_{t}-z_{t-1}){\frac {|z_{t}-
z_{t-1}|^{N}}{2r_{\nu }\mu _{\nu }^{N}}},\nu =\nu (x_{t})=\nu (x_{t-1}),
\end{displaymath}
and then increment \(k\) by 1.

@@

The conditions of convergence of the described in Section \ref{sec:method} method in case of \(q=1\) are given in \cite{Strongin2000}.
\begin{theorem} (Sufficient convergence conditions)
  \label{th:single_conv}
  Assume that the following conditions are true:
  \begin{enumerate}
    \item The problem (\ref{eq:constrained_problem}) has a solution.
    \item Functions \(g_j(y)\leqslant 0, 1\leqslant j\leqslant m + 1\), are Lipschitzian with
respective constants \(L_i\) over the domain \(D\) (here \(g_{m+1}(y)=\varphi(y)\)).
    \item Beginning with some step (i.e. if \(k\) from (\ref{eq:points}) is sufficiently large),
    the values \(\mu_\nu\) from (\ref{step2}) satisfy the inequalities
    \begin{equation}
      r_\nu\mu_\nu > 2^{3-1/N}L_\nu \sqrt{N+3},\: 1\leqslant \nu \leqslant m + 1.
    \end{equation}
  \end{enumerate}
  Then any limit point \(\overline{y}\) of the sequence \(\{y_k\} = \{y(x_k)\}\) generated
  by the index method while solving the problerm
  (\ref{eq:constrained_problem}) is admissible and satisfies the conditions
\begin{equation}
  \label{eq:conv_cond}
  \varphi(\overline{y})=\{ \varphi(y^k): g_i(y^k)\leqslant 0,1\leqslant j\leqslant m, k=1,2,\dots\}=\varphi(y^*),
\end{equation}
  where \(y^*\) is from (\ref{eq:constrained_problem}).
\end{theorem}

@@

\begin{theorem} (On convergence conditions) Let the conditions of the Theorem \ref{th:single_conv}
  are satisfied for all of \(q\) problems from the set (\ref{eq:many_problems}) separately.
  Then solving all of the \(q\) problems by the algorithm from Section \ref{sec:method} will
  generate \(q\) infinite sequences, for the limit points of which the Theorem \ref{th:single_conv}
  will be also true.
\end{theorem}
\begin{proof}
  Let us consider two random problems from the set (\ref{eq:many_problems})
  \begin{equation}
      \begin{array}{lr}
        \min\{\varphi(y):y\in D_1,\: g_j^\varphi(y)\leqslant 0, 1\leqslant j\leqslant m_1\}, \\
        \min\{\psi(y):y\in D_2,\: g_j^\psi(y)\leqslant 0, 1\leqslant j\leqslant m_2\}.
      \end{array}
  \end{equation}
  Let us denote the characteristics (\ref{step3_1}) for the first problem as \(R_\varphi(i)\)
  and for the second problem as \(R_\psi(j)\). Considering that, we have:
  \begin{equation}
      \begin{array}{lr}
        R_\varphi(t_\varphi)=\max_{1\leqslant i\leqslant k}R_\varphi(i), \\
        R_\psi(t_\psi)=\max_{1\leqslant j\leqslant s}R_\psi(j),
      \end{array}
  \end{equation}
  where \(k\) corresponds to the number of trials in the first problem and \(s\)
  corresponds to the number of trials in the second one. The sequence of trials \(\{v^k\}\)
  corresponds to the first problem and the sequence of trials \(\{u^s\}\)
  corresponds to the second problem. The values \(z^k=g^\varphi_\nu(v_k),\nu =\nu (v_{k})\) correspond to the
  trial points \(\{v^k\}\) and the values \(w^s=g^\psi_\nu(u_s),\nu =\nu(u_{s})\) correspond to the
  trial points \(\{u^s\}\).

@@

  Let us consider the first case: from (\ref{step2})
  \begin{displaymath}
    \frac{|z_i-z_{i-1}|}{\Delta_i} \leqslant \mu_{\varphi,\nu}.
  \end{displaymath}
  Considering that, we can establish an upper bound:
  \begin{displaymath}
    \frac{(z_i-z_{i-1})^2}{r_\nu^2\mu_{\varphi,\nu}^2\Delta_i}=\frac{(z_i-z_{i-1})^2\Delta_i}{(r_\nu\mu_{\varphi,\nu}\Delta_i)^2}
    \leqslant \frac{\Delta_i}{r_\nu^2}.
  \end{displaymath}
  Therefore using the first branch form rule (\ref{step3_1}) we have inequality
  \begin{equation}
    \label{eq:th1}
    R_\varphi(i)\leqslant\Delta_i(1 + \frac{1}{r_\nu^2}) - \frac{2(z_i+z_{i-1}-2z^*_\nu)}{r_\nu\mu_{\varphi,\nu}}.
  \end{equation}
  Because \(\overline{v}\) is the limit point of the sequence \(\{v^k\}\) and \(\varphi(y(\overline{v}))\leqslant z^*_{\nu}, \nu=m+1\) or
  \(z^*_\nu=0, \nu<m+1\) and \(z_{i-1},z_i\to 0\) at \(k\to\infty\):
  \begin{equation}
    \label{eq:th2}
    \Delta_i\to 0, z_i+z_{i-1} - 2 z_\nu^*\to 0.
  \end{equation}
  In the second case (when one of the other two branches of the rule (\ref{step3_1}) is applied) we have
  \begin{displaymath}
    R_\varphi(i)=2\Delta_i - 4\frac{z_i-z^*_\nu}{r_\nu\mu_{\varphi,\nu}}.
  \end{displaymath}
  If \(z^*_\nu \ne 0\) then \(z_i-z^*_\nu \geqslant 0\) and
  \begin{equation}
    \label{eq:th3}
    R_\varphi(i)=2\Delta_i - 4\frac{z_i-z^*_\nu}{r_\nu\mu_{\varphi,\nu}} \leqslant 2\Delta_i.
  \end{equation}
  Otherwise because of \(\overline{v}\) is a feasible point \(z_i\to 0\) at \(k\to\infty\).

@@

  If \(\nu(w_{j-1})=\nu(w_{j})=\alpha\) then the first branch of the rule (\ref{step3_1}) is applied and
  \begin{dmath}
    R_\psi(j)=\Delta_j + \frac{(w_j-w_{j-1})^2}{r_\alpha^2\mu_{\psi,\alpha}^2\Delta_j}
      - \frac{2(w_j+w_{j-1}-2w^*_\alpha)}{r_\alpha\mu_{\psi,\alpha}} \geqslant
      \Delta_j - \frac{2(w_j+w_{j-1}-2w^*_\alpha)}{r_\alpha\mu_{\psi,\alpha}} =
      \Delta_j - \frac{2\Delta_j(w_{j-1}-w_j)}{r_\alpha\mu_{\psi,\alpha}\Delta_j} \geqslant
      \Delta_j - \frac{\Delta_j}{r_\alpha} = \Delta_j\left(1-\frac{1}{r_\alpha}\right).
    \label{eq:th4}
  \end{dmath}
